Drivers of tree regeneration remain challenging to be captured by models of forest dynamics

### 2022-11-23 manuscript/manuscriptRW.Rmd
keep all the methodological limitation in the discussion with an expansion in other parts

---
title: "Drivers of tree regeneration remain challenging to be captured by models of forest dynamics"
author: 
- Olalla Díaz-Yáñez, 
- Yannek Käber, 
- Mats Mahnken, 
- Samuel Fischer, 
- Xavier Morin, 
- Manfred Lexer, 
- Christian Hochauer, 
- Katarina Merganicova, 
- Tobias Mette, 
- Werner Rammer, 
- Kristin H. Braziunas, 
- Paola Mairota, 
- Josef Bruna,
- Heike Lischke, 
- Daniel Scherrer, 
- Simon Scheiter, 
- Thomas Hickler, 
- Tim Anders,
- Jessica Hetzer, 
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  word_document:
    reference_docx: template_paper.docx
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Regeneration is a key process in forest dynamics. Models of forest
dynamics have been long lagging an assessment in their capacities to
properly represent this process. Using 15 models and variants of forest
dynamics at the stand, landscape and global levels, we simulate the
regeneration across 200 sites representing environmental gradients
across Europe. The simulation results were compared with observed data
collected from forest reserves across Europe. In this study we want to
asses how well are models of forest dynamics capturing regeneration
levels, species composition and mortality in regeneration. The results
show that .

@@

Tree recruitment depends on multiple processes such as reproduction,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival; these processes are all difficult to assess and
some of them are scarcely understood and thus appear highly stochastic
[@bogdziewicz2021]. Finding the right level of complexity in the
definition of the factors leading to tree recruitment is challenging and
models of forest dynamics have used different approaches [@bugmann2022].
Overall, models are needed to (1) synthesize existing empirical data and
explore their consequences, and (2) assess future tree recruitment e.g.
in the context of global change scenarios. Currently, regeneration
processes in dynamic forest models are handled in a multitude of ways
[@konig_tree_2022; @bugmann2022], from entirely ignoring tree
regeneration (as done in classical forest growth models, e.g.
@pretzsch2002), across the use of a few simple environmental filters, as
done in most forest \"gap\" models (e.g., @shuman2011) and Dynamic
Global Vegetation Models (e.g., @smith2001; @hickler2012), to complex
approaches that incorporate local feedback from the canopy, multiple
ecological processes and often also small time steps (e.g., @wehrli2006;
@seidl2012), or field-based statistical parameterizations, which however
are not easy to extrapolate in space and time (e.g., @ribbens1994).

With the current strategies that models of forest dynamics are using to
represent tree recruitment we are obtaining results that faced several
problems and have strong impacts on forest dynamic such as
over-recruitment, needed excess of mortality at early stages and
inadecuate species composition in regeneration. Some model outputs are
over-recruiting which subsequently must die primarily by competition in
later stages. This is characteristic of \'classical\' forest gap models
[@gutiérrez2016], at least as long as the simulations comprise a
multitude of species. Other models use a model output to correct
recruitment levels for example by routinely calibrate them against the
predicted canopy species composition affecting as well the species
dominance.

@@

Based on their establishment modules the models can be categorized as
regeneration or recruitment (or ingrowth) models [@vanclay1994;
@konig_tree_2022] (see Table \@ref(tab:tableModels)). Regeneration
models include processes such as flowering and pollination, seed
production, seed dispersal, germination or seedling establishment that
lead to the number of seedlings or saplings. Recruitment models
introduce a number of seedlings or saplings with certain characteristic
such as biomass or dbh, without considering earlier development
processes. We can also distinguish models that have feedback in the
regeneration as those in which the level and species composition in
regeneration is influenced by the existing adult tree community via the
production of seeds, seedlings or saplings. The starting point for
regeneration it is also different per model varying from seedbank, seed,
sapling or ingrowth. The overall complexity in the establishment modules
also varies, following @bugmann2022 we can see that most of the models
included in this study can be classified as having rather low complexity
in their tree recruitment modules while few of them have high complexity
specifically when defining processes such as XXXX, XXXX (enumerate
models). Further details on the models specific formulation related to
their regeneration modules and an overall summary across models is
presented in the supplementary materials (SHALL WE DO THIS?).

@@

kable(modelDF, 
      #format = "markdown",
      format = "latex", booktabs = TRUE,
      caption = "List of models included.") |>
        kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
    # fix width by column
      kableExtra::column_spec(column = 1, width = "5em") |> 
      kableExtra::column_spec(column = 2, width = "5em") |> 
      kableExtra::column_spec(column = 3, width = "5em") |> 
      kableExtra::column_spec(column = 4, width = "5em") |>
      kableExtra::column_spec(column = 5, width = "5em") |>
      kableExtra::column_spec(column = 6, width = "5em") |>
      kableExtra::column_spec(column = 7, width = "8em") |>
      kableExtra::column_spec(column = 8, width = "5em") |>
      kableExtra::column_spec(column = 9, width = "5em") |>
      kableExtra::column_spec(column = 10, width = "5em") |>
      kableExtra::column_spec(column = 11, width = "5em") |>
      kableExtra::column_spec(column = 12, width = "5em") 
```

@@

Each modeler group was provided with detailed protocol (available in the
supplementary material) with instruction on how to perform the
simulations regarding input variables such as climate or soil,
simulation setup and expected output variables. No information was
provided on the regeneration prior to the simulation. The overarching
goal of this experiment was the assessment of the regeneration niche as
it arises from empirical data analysis against its representation in a
wide range of models of forest dynamics.

@@

Each of the models reported the number of recruits over a 10 year
interval per species and per ha for each of the sites. To better
understand the variation per site each modeling team decided on how to
sample 200 multiple values per site to show the variation in recruitment
for spatial units of 1 ha during the simulated phase of potential
natural vegetation but without providing average of multiple 1-ha
samples. This was done with different strategies such as sampling
simulated data from the same 1-ha plot in the equilibrium over time, or
by sampling several 1-ha plots from the simulated forest at one specific
point in time (in the equilibrium). This resulted in 880000 observation
per model that simulated the 11 species indicated in the protocol (200
sites, 200 samples per site, 11 species and two diameter thresholds).
Those models that simulated extra species, the recruitment of those
extra species were aggregated in a category others. The simulated data
per model and the observations used in this study are provided as a
supplementary information OR IN A REPOSITORY? available here(XXX).

@@

```{r nvTrend710BA, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites against the stand basal area. (A) y-axis scaled by model; (B) same y-axis across all models. The values were split into 10 bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
    knitr::include_graphics(c(here::here("figures", "envTrend7Totba.png"),
                              here::here("figures", "envTrend7Totba_500.png")))
```

@@

Most of the models match the level of diversity (as measured by the
Shannon index) of the observed data quite well (Figure \@ref(fig:H710)),
although they consistently underestimate its variability across sites.
Four models overestimate recruitment diversity: 4C, PICUS, LANDIS-II and
LPJ-GUESS, the latter mostly for the 10 cm diameter threshold. The model
4C is a special case as this model simulated 5 species only. In most
models there are small differences across sites regarding the species
richness of recruitment between 7 and 10 cm. Most models maintain the
recruitment diversity between the 7 and 10 cm thresholds, and the same
is visible from the observed data, there are only few models where the
recruitment diversity increases between 7 and 10 cm which might be an
artifact of the sample strategy and calculating the Shannon index based
in basal area as for a same sample recruitment at 10cm will include
larger trees with larger basal area and if those are from different
species, the Shannon index will be larger.

@@

```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site and for observed and simulated values. A) Shows the three trends present across models for the species richness at the stand level. B) Shows the three trends present across models for recruitment at 7cm threshold. The selected representative model per trend group was based on supplementary figure 5 and 6", out.width='75%'}
knitr::include_graphics(here::here("figures", "combined_richness.png"))
```

@@

```{r diff, echo=FALSE, fig.cap="Departure of each model in the representation of the main species recruitment relative abundance from the observations across the environmental gradients for a 7cm treshold. The values shown are the mean of samples per site and across sites in bins, where the range of the climatic water balance (cm) and seasonal degree-day sum (dºC) was divided in 10 bins. The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species. The color gradient shows the difference between the simulated and observed ratio between recruited basal area of the species and the total recruited basal area.", out.width='100%', out.height= '90%'}
knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
```

@@

There are strong differences across models on the mortality rate between
7 and 10 cm (Figure \@ref(fig:mortality11)). All models except for the
two DGVMs LPJ-GUESS and aDGVM2 feature a mortality rate larger than the
observed data, thus at least partially compensating for the general
overestimation of recruitment rates (Figure \@ref(fig:over)). Some of
the models showed strong compensation for the initial overestimation of
recruitment with larger mortality of recruitment between 7 and 10 cm
thresholds but not in others. On the one hand, models below the Reineke
line have lower mortality than expected by the self-thinning rule, which
may point to facilitation, or simply a higher mortality rate before the
trees reached 7 cm diameter. On the other hand, models such as xComp,
PICUS and TreeMig are above the self-thinning line (Reineke), indicating
that they have higher mortality in recruitment of mixed species forest
than the expected in an even-aged single species forests, which
indicates strong compensation for the initial overestimation of
recruitment in some models. However, some of the models simulated a
mortality rate in recruitment that is broadly compatible with the
self-thinning rule, i.e. 4C, SIBYLA , LandClim and LANDIS II.

@@

Tree recruitment is a fundamental process in forest dynamics, in
particular with increasing disturbance under climate change . Correctly
capturing regeneration in forest dynamics is fundamental to evaluate for
example post-disturbance dynamics as this will define the status of the
forests in the first 15-20 years [@seidl_post-disturbance_2022]. If this
initial state it is not well captured then we can not properly assess
aspects such as the forest resilience, or changes in carbon capture. In
this study, for the first time, projections of tree recruitment done by
multitude models of forest dynamics were confronted with observed data
across a large environmental gradients. The observed data that we used
[@käber2022] is exceptional in this regard, and that's why it was used
for model benchmarking rather than calibration.

Adequate tree recruitment rates define the structure and composition of
forests. Most of the models of forest dynamics tended to over-recruit
trees in all the sites. Excess in recruitment has to lead to a large
mortality in later stages. Capture properly recruitment rates in the
environmental gradient might impact future forests. If not capture well
it is going to be difficult to for models to correctly project range
shifts retractions in species expansions because of climate change might
not also be properly represented if we do not capture properly
recruitment rates in the environmental gradient.

Regeneration levels define the success of future forests, but it is
equally important to correctly identify un-stocked forest areas
[@rammer2021]. Tree recruitment is often characterized by an excess
number of zero counts [@fortin2007]. In this study some of the models
did not have lack of recruitment or their proportion was very low. This
substantial difference with the observed data might be a consequence of
assuming that the observed data is also in equilibrium, where lack of
recruitment in sites could also be an indication of "non-equilibrium"
due to management history.

Light availability is an important process defining tree recruitment
[@collet2006; @berdanier2016]. However it is difficult to have measured
variables defining light availability for the regeneration. In this
study we used the total stand basal area as a proxy for light
availability. There was differences in the basal area ranges, some of
the models had stand basal area lower than the observe data but there
was also models that did have more stocked stands that the observed data
and therefore it was not possible to evaluate the recruitment for lower
density stands. It was also not possible to observed the expected
pattern showing how recruitment of the different species is becoming
sparser and drops out completely as a function of a proxy of light
availability.

If the recruitment tree composition is not well captured then we can not
assess the functional composition of the forest appropriately, e.g.
their sensitive to drought or resilience to new disturbances events. In
the models simulation we saw variations on how the main tree species
were represented in the environmental gradient but also in the in
overall species richness levels. Most models maintain or even increase
in recruitment diversity between the 7 and 10 cm thresholds, and the
same is visible from the observed data, however the differences are not
significant. Species richness variations might be explained by the way
each model considers recruitment, e.g. in the absence of adults
recruitment can be linked to either seed rain from elsewhere -dispersal,
or to seed/seedling bank). Differences might also be expected between
cohort and individual trees based models.

-   Why models with feedback also underestimate species richness in
    recruitment?

-   Is the adult forest the right forest, how is this relevant and is
    this the consequence for the correct recruitment composition? This
    is highly dependent on the model structure:Background seed Yes/no /
    Stochastic vs deterministic

-   Why the models over predicting species richness at the stand level
    are all the "ForClim family" and FORMIND ?

We could expect that models that over-recruited trees. will also have a
large mortality between 7 and 10 cm thresholds. This implies that this
models have to have the excess of mortality in later stages of the
forest dynamics. Models below the Reineke line have lower mortality than
expected by the self-thinning rule, which may point to facilitation, or
simply a higher mortality rate before the trees reached 7 cm diameter,
however one would also assume that these models will underestimate
recruitment which was not the case. Why?. There might be a difference
between models including stochastic processes in recruitment and those
which do not, but perhaps the reality implies higher mortality rates in
the adult component (enhance the share of less shade tolerant species).

There is also the question of how much level of detail one have to give
to define regeneration processes [@konig_tree_2022; @bugmann2022].
Models with more detailed and complex recruitment modules, will have
also increase parametrisation which will lead to lower generalization as
this values will have to be collected from specific geographical
regions.

-   how it relates to the results

-   future model development recommendations

**Methodological aspects**

@@

-   Recruitment of trees into size classes of ≥4 cm DBH, as in our four
    data sets, represents a relatively late stage of forest
    regeneration. It integrates over sometimes long periods since
    germination and multiple environmental constraints act upon young
    trees of widely different size. kaber 2021

@@

>>>>

---
title: "Drivers of tree regeneration remain challenging to be captured by models of forest dynamics"
author: 
- Olalla Díaz-Yáñez, 
- Yannek Käber, 
- Mats Mahnken, 
- Samuel Fischer, 
- Xavier Morin, 
- Manfred Lexer, 
- Christian Hochauer, 
- Katarina Merganicova, 
- Tobias Mette, 
- Werner Rammer, 
- Kristin H. Braziunas, 
- Paola Mairota, 
- Josef Bruna,
- Heike Lischke, 
- Daniel Scherrer, 
- Simon Scheiter, 
- Thomas Hickler, 
- Tim Anders,
- Jessica Hetzer, 
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Regeneration is a key process in forest dynamics. Models of forest
dynamics have been long lagging an assessment in their capacities to
properly represent this process. Using 15 models and variants of forest
dynamics at the stand, landscape and global levels, we simulate the
regeneration across 200 sites representing environmental gradients
across Europe. The simulation results were compared with observed data
collected from forest reserves across Europe. In this study we want to
asses how well are models of forest dynamics capturing regeneration
levels, species composition and mortality in regeneration. The results
show that most of the models of forest dynamics tend to over-estimate
recruitment rates although this is not always connected to large
mortality rates in early dynamics stages. The effect of light
availability on recruitment rates is better captured than temperature or
soil moisture, however the patterns are not consistent across models.
Species richness matches quite well compared to the ranges observed in
the data. Models simulating higher species diversity at the stand level
do not feature higher diversity in recruitment. *Abies alba* was well
captured across temperature and soil moisture gradients by most models
and *Fagus sylvatica* was generally underestimated. *Picea abies*
recruitment was generally over-estimated on colder sites. HERE TWO
SENCENTECES DISCUSSION/CONCLUSION. HERE A SO WHAT SENTENCE

@@

Tree recruitment depends on multiple processes such as reproduction,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival; these processes are all difficult to assess and
some of them are scarcely understood and thus appear highly stochastic
[@bogdziewicz2021]. Finding the right level of complexity in the
definition of the factors leading to tree recruitment is challenging and
models of forest dynamics have used different approaches [@bugmann2022].
Overall, models are needed to (1) synthesize existing empirical data and
explore their consequences, and (2) assess future tree recruitment e.g.
in the context of global change scenarios. Currently, regeneration
processes in dynamic forest models are handled in a multitude of ways
[@konig_tree_2022; @bugmann2022], from entirely ignoring tree
regeneration (as done in classical forest growth models, e.g.
@pretzsch2002), across the use of a few simple environmental filters, as
done in most forest "gap" models (e.g., @shuman2011) and Dynamic Global
Vegetation Models (e.g., @smith2001; @hickler2012), to complex
approaches that incorporate local feedback from the canopy, multiple
ecological processes and often also small time steps (e.g., @wehrli2006;
@seidl2012), or field-based statistical parameterizations, which however
are not easy to extrapolate in space and time (e.g., @ribbens1994).

With the current strategies that models of forest dynamics are using to
represent tree recruitment we are obtaining results that faced several
problems and have strong impacts on forest dynamic such as
over-recruitment, needed excess of mortality at early stages and
inadecuate species composition in regeneration. Some model outputs are
over-recruiting which subsequently must die primarily by competition in
later stages. This is characteristic of 'classical' forest gap models
[@gutiérrez2016], at least as long as the simulations comprise a
multitude of species. Other models use a model output to correct
recruitment levels for example by routinely calibrate them against the
predicted canopy species composition affecting as well the species
dominance.

@@

Based on their establishment modules the models can be categorized as
regeneration or recruitment (or ingrowth) models [@vanclay1994;
@konig_tree_2022] (see Table \@ref(tab:tableModels)). Regeneration
models include processes such as flowering and pollination, seed
production, seed dispersal, germination or seedling establishment that
lead to the number of seedlings or saplings. Recruitment models
introduce a number of seedlings or saplings with certain characteristic
such as biomass or dbh, without considering earlier development
processes. We can also distinguish models that have feedback in the
regeneration as those in which the level and species composition in
regeneration is influenced by the existing adult tree community via the
production of seeds, seedlings or saplings. The starting point for
regeneration it is also different per model varying from seedbank, seed,
sapling or ingrowth. The overall complexity in the establishment modules
also varies, following @bugmann2022 we can see that most of the models
included in this study can be classified as having rather low complexity
in their tree recruitment modules while few of them have high complexity
specifically when defining processes such as XXXX, XXXX (enumerate
models).

@@

kable(modelDF, 
      #format = "markdown",
      format = "latex", booktabs = TRUE,
      caption = "List of models included.") |>
        kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
    # fix width by column
      kableExtra::column_spec(column = 1, width = "5em") |> 
      kableExtra::column_spec(column = 2, width = "5em") |> 
      kableExtra::column_spec(column = 3, width = "5em") |> 
      kableExtra::column_spec(column = 4, width = "5em") |>
      kableExtra::column_spec(column = 5, width = "5em") |>
      kableExtra::column_spec(column = 6, width = "5em") |>
      kableExtra::column_spec(column = 7, width = "5em") |>
      kableExtra::column_spec(column = 8, width = "8em") |>
      kableExtra::column_spec(column = 9, width = "8em") |>
      kableExtra::column_spec(column = 10, width = "5em") |>
      kableExtra::column_spec(column = 11, width = "5em") |>
      kableExtra::column_spec(column = 12, width = "5em") 
```

@@

Each modeler group was provided with a detailed protocol (available in
the supplementary material) with instruction on how to perform the
simulations regarding input variables such as climate or soil,
simulation setup and expected output variables. No information was
provided on the regeneration prior to the simulation. The overarching
goal of this experiment was the assessment of the regeneration niche as
it arises from empirical data analysis against its representation in a
wide range of models of forest dynamics.

@@

Each of the models reported the number of recruits over a 10 year
interval per species and per ha for each of the sites. To better
understand the variation per site each modeling team decided on how to
sample 200 multiple values per site to show the variation in recruitment
for spatial units of 1 ha during the simulated phase of potential
natural vegetation but without providing average of multiple 1-ha
samples. This was done with different strategies such as sampling
simulated data from the same 1-ha plot in the equilibrium over time, or
by sampling several 1-ha plots from the simulated forest at one specific
point in time (in the equilibrium). This resulted in 880000 observation
per model that simulated the 11 species indicated in the protocol (200
sites, 200 samples per site, 11 species and two diameter thresholds).
Those models that simulated extra species, the recruitment of those
extra species were aggregated in a category others. The simulated data
per model and the observations used in this study are provided as a
supplementary information / repository.

@@

```{r nvTrend710BA, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites against the stand basal area. (A)same y-axis across all models; (B)y-axis scaled by model. The values were split into 10 bins; the red lines represent a GAM model showing the trend in the observed data.", fig.show = 'hold', out.width = '80%'}
    knitr::include_graphics(c(here::here("figures", "envTrend7Totba_500.png"),
                              here::here("figures", "envTrend7Totba.png")))
```

@@

Most of the models match the level of diversity (as measured by the
Shannon index) of the observed data quite well (Figure \@ref(fig:H710)),
although they consistently underestimate its variability across sites.
Four models overestimate recruitment diversity: 4C, PICUS, LANDIS-II and
LPJ-GUESS, the latter mostly for the 10 cm diameter threshold. The model
4C is a special case as this model simulated 5 species only. In most
models there are small differences across sites regarding the species
richness of recruitment between 7 and 10 cm. Most models maintain the
recruitment diversity between the 7 and 10 cm thresholds, and the same
is visible from the observed data, there are only few models where the
recruitment diversity increases between 7 and 10 cm which might be an
artifact of the sample strategy and calculating the Shannon index based
in basal area as for a same sample recruitment at 10 cm will include
larger trees with larger basal area and if those are from different
species, the Shannon index will be larger.

@@

```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site and for observed and simulated values. A) Shows the three trends present across models for the species richness at the stand level. B) Shows the three trends present across models for recruitment at 7 cm threshold. The selected representative model per trend group was based on the supplementary figure 5 and 6.", out.width='75%'}
knitr::include_graphics(here::here("figures", "combined_richness.png"))
```

@@

```{r diff, echo=FALSE, fig.cap="Departure of each model in the representation of the main species recruitment relative abundance from the observations across the environmental gradients for a 7cm treshold. The values shown are the mean of samples per site and across sites in bins, where the range of the climatic water balance (cm) and seasonal degree-day sum (dºC) was divided in 10 bins. The point sizes represent ratio between recruited basal area of the species and the total recruited basal area for all the species, where a lack of dot represents zero basal area recruitment share. The color gradient shows the difference between the simulated and observed ratio of recruited basal area of the species and the total recruited basal area for all the species.", out.width='100%', out.height= '90%'}
knitr::include_graphics(here::here("figures", "HeatEnvGrad_10circles_7.png"))
```

@@

There are strong differences across models on the mortality rate between
7 and 10 cm (Figure \@ref(fig:mortality11)). All models except for the
two DGVMs LPJ-GUESS and aDGVM2 feature a mortality rate larger than the
observed data, thus at least partially compensating for the general
overestimation of recruitment rates (Figure \@ref(fig:over)). Some of
the models showed strong compensation for the initial overestimation of
recruitment with larger mortality of recruitment between 7 and 10 cm
thresholds but not in others. On the one hand, models below the Reineke
line have lower mortality than expected by the self-thinning rule, which
may point to facilitation, or simply a higher mortality rate before the
trees reached 7 cm diameter. On the other hand, models such as xComp,
PICUS and TreeMig are above the self-thinning line (Reineke), indicating
that they have higher mortality in recruitment of mixed species forest
than the expected in an even-aged single species forests, which
indicates strong compensation for the initial overestimation of
recruitment in some models. However, some of the models simulated a
mortality rate in recruitment that is broadly compatible with the
self-thinning rule, i.e. 4C, SIBYLA, LandClim and LANDIS II.

@@

Tree recruitment is a fundamental process in forest dynamics, in
particular with increasing disturbance under climate change. Correctly
capturing regeneration in forest dynamics is fundamental to for example
evaluate post-disturbance dynamics as this will define the status of the
forests in the first 15-20 years [@seidl_post-disturbance_2022]. If this
initial state it is not well captured then we can not properly assess
aspects such as the forest resilience, or changes in carbon capture. In
this study, for the first time, projections of tree recruitment done by
several models of forest dynamics were confronted with observed data
across a large environmental gradients. The observed data that we used
[@käber2022] is exceptional in this regard, and to use it for model
benchmarking has gave us much more insights than if we have used it for
calibration.

Adequate tree recruitment rates define the structure and composition of
forests. Most of the models of forest dynamics tended to over-recruit
trees in all the sites. Over estimating recruitment might have
fundamental implications on the role of regrowth and forest carbon, with
an overestimation in their capacity for carbon capture for example after
the impact of disturbances [@pugh2019]. Excess in recruitment also
requires of larger mortality rates in later stages, changing the
expected forest densities at early stages and forcing the models to have
stronger filters. Regeneration levels define the success of future
forests, but it is equally important to correctly identify un-stocked
forest areas [@rammer2021]. Tree recruitment is often characterized by
an excess number of zero counts [@fortin2007]. In this study some of the
models did not have lack of recruitment or their proportion was very
low. This substantial difference with the observed data might be a
consequence of assuming that the observed data is also in equilibrium,
where lack of recruitment in sites could also be an indication of
"non-equilibrium" due to management history.

Light availability effect on tree regeneration have been widely
documented [@messier1999; @collet2006; @berdanier2016]. However it is
difficult to have variables defining light availability for the
regeneration. In this study, we used the total stand basal area as a
proxy for light availability. There was differences in the basal area
ranges, some of the models had stand basal areas lower than the observed
data but there was also models that did have more stocked stands than
the observed data, therefore it was not possible to evaluate the
recruitment for lower density stands. Regeneration at low stand
densities is an interesting aspect to evaluate how well forests are
recovering e.g. after the creation of gaps due to disturbances. We would
have expected that the recruitment of the different species would become
sparser and droped out completely as a function of a proxy of light
availability, however few models showed this trend indicating that the
relation between recruitment and light availability it is not yet
correctly defined.

When tree composition in recruitment is not well captured, then we can
not assess the functional composition of the forest appropriately, e.g.
their sensitive to drought or resilience to new disturbances events
[@redmond2015; @seidl_post-disturbance_2022]. In the models simulation
outputs overall species richness levels in recruitment were inside the
observed ranges, however there were variations on how the main tree
species were represented in the environmental gradient. Most models
maintain or even increase in recruitment diversity between the 7 and 10
cm thresholds, and the same is visible from the observed data, however
the differences are not significant. Species richness variations might
be explained by the way each model considers recruitment, e.g. in the
absence of adults, recruitment can be linked to either seed
rain-dispersal, or to seed/seedling bank, but also on what processes are
formulated with stochastic or deterministic approach. Differences might
also be expected between cohort and individual trees based models
**how?**. However, our results do not show differences between models
with an without feedback in recruitment, **why?** The adult forests
species richness was better captured than the recruitment species
richness, and only few models over-predicted stand level species
richness **why these ones?**.

We could expect that models that over-recruit trees will also have a
large mortality between 7 and 10 cm thresholds but this was not shown.
This implies that these models must have the excess of mortality in
later stages. Higher mortality in the adult component might enhance the
share of less shade tolerant species for those models. There might be a
difference between models including stochastic processes in recruitment
and those which do not ??. There was also a group of models below the
Reineke line indicating a lower mortality than expected by the
self-thinning rule, which may point to facilitation, or simply a higher
mortality rate before the trees reached 7 cm diameter, another
possibility would have been that these models underestimate recruitment
which was not the case.

There is also the question of how much level of detail one have to give
to define regeneration processes [@konig_tree_2022; @bugmann2022].
Models with more detailed and complex recruitment modules, will have
also increase parametrisation which will lead to lower generalization as
this values will have to be collected from specific geographical
regions. This is significant relevant when looking at mixed forests and
the species level parametrization. **How it relates to the results?
Future model development recommendations?**

@@

-   Recruitment of trees into size classes of larger 4 cm DBH, as in our
    four data sets, represents a relatively late stage of forest
    regeneration. It integrates over sometimes long periods since
    germination and multiple environmental constraints act upon young
    trees of widely different size. kaber 2021

@@

**Conclusion paragraph**

Capture properly recruitment rates in the environmental gradient might
impact future forests.
